---
title: "About"
toc: false
---

```{r}
#| label: table-setup
#| echo: false
#| message: false

library("tidyverse")
library("kableExtra")
```

In this course students examine the fundamental principles of  doing experimental research in linguistics. 
Specifically, the focus is on developing an in depth understanding of the experimental paradigms and statistical procedures used in sociolinguistics, phonetics, psycholinguistics, syntax, and corpus linguistics. 
Students will learn advanced techniques used to explore, tidy, visualize, and analyze data. 
We will also focus on how to make the aforementioned procedures reproducible and shareable. 
Students will develop a foundation in programming in R, as well as learning the most common tools at the disposal of todays data scientist (i.e. GitHub, Knitr, etc.). 
No prior experience with statistics or programming is necessary.


## Course info

```{r}
#| label: info
#| echo: false

home <- '<a href="http://www.jvcasillas.com/" title="homepage"><i class="fa fa-home"></i></a>'
mail <- '<a href="mailto:joseph.casillas@rutgers.edu"><i class="fa fa-paper-plane fa-fw"></i></a>'
github <- '<a href="http://github.com/jvcasillas"><i class="fa fa-github fa-fw"></i></a>'
twitter <- '<a href="http://twitter.com/jvcasill"><i class="fa fa-twitter fa-fw"></i></a>'

tribble(
  ~" ", ~"  ", 
  "Instructor:", paste("Joseph Casillas (", home, mail, github, twitter, ")"), 
  "When:", "Thursday 08:10 - 11:10am", 
  "Where:", "Online", 
  "Office hours:", "By appointment"
) %>% 
  knitr::kable(format = "pandoc") 
```



## Materials

```{r}
#| label: materials
#| echo: false

r4ds <- "Wickham, H., & Grolemund, G. (2017). R for Data Science. O'Reilly Media."
johnson <- "Johnson, K. (2008). Quantitative Methods in Linguistics. Blackwell Publishing."
qass22 <- "Lewis-Beck, M. (1980). Applied Regression: An Introduction. Volume 22 of Quantitative Applications in the Social Sciences. Sage Publications."
qass50 <- "Berry, W., & Feldman, S. (1985). Multiple Regression in Practice. Issue 50. Quantitative Applications in the Social Sciences. Sage Publications."
qass57 <- "Schroeder, L. D., Sjoquist, D. L., & Stephan, P. E. (1986). Understanding Regression Analysis: An Introductory Guide. Issue 57. Quantitative Applications in the Social Sciences. Sage Publications."
qass93 <- "Hardy, M. A. (1993). Regression with Dummy Variables. Issue 93. Quantitative Applications in the Social Sciences. Sage Publications."

links <- c(
  "http://r4ds.had.co.nz", 
  "http://a.co/9ArHNqO", 
  "http://a.co/2zyfhXj", 
  "http://a.co/8jcU0ef", 
  "http://a.co/gE2xmLr", 
  "http://a.co/8Spewd9"
)


tribble(
  ~"item", ~"ref", ~"link",
  1, r4ds,    "Online", 
  2, johnson, "Amazon", 
  3, qass22,  "Amazon", 
  4, qass50,  "Amazon", 
  5, qass57,  "Amazon", 
  6, qass93,  "Amazon"
) %>% 
  kbl(align = c("l", "l", "r")) %>% 
  row_spec(0, color = "white") %>% 
  column_spec(3, link = links, width = "10em") %>% 
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive")
  )
```


<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">
