---
title             : "Effects of perceptual training on ability to produce L2 English plosives"
shorttitle        : "Production of L2 plosives"

author: 
  - name          : "Ana Noelle Bennett"
    corresponding : yes    # Define only one corresponding author
    address       : "1498 State Route 28 West Hurely, NY 12491"  
    email         : "anb136@psych.rutgers.edu"  


affiliation:
    institution   : "Rutgers, the State University of New Jersey"


author_note: | 
 Cognitive Psychology   
 Center for Cognitive Science 

abstract: |
  Enter abstract here. Each new line herein must be indented, like this line.
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no

class             : "man"
output            : papaja::apa6_pdf
---

```{r load_packages, include = FALSE}
library("papaja")
```

```{r analysis_preferences}
# Seed for random number generation
set.seed(42)
```



# Methods
In this study, we conducted acoustic analyses on Cantonese speakers' productions of English phonological minimal word pairs with voiced (i.e. /b d g/) and voiceless (i.e. /p t k/) plosives in coda position. Specifically, we utilized PRAAT to measure the duration of the vowel. This was motivated by evidence that vowel duration is an acoustic cue that indicates in English if the following plosive is voiced or voiceless  -- with the duration of the vowel preceding a voiced stop being longer than the duration of a vowel preceding a voiceless stop (Charles-Luce, 1985; House and Fairbanks, 1953; Peterson and Lehiste, 1960; House, 1961; Umeda, 1975; Klatt, 1976). The production of the word "got" was excluded from this analysis as it was the only word that did not have a minimal pair.Please note that the production data analyzed in this study was collected and generously provided by Dr. Terry Kit-fong Au, from the University of Hong Kong. 

## Participants
There were a total of 36 University students from the University of Hong Kong. 18 of the participants were in the training group (33% men), and 18 of the participants were in a wait-list control group (28% men). 

## Material
The following analyses are based on productions of phonological minimal word pairs with voiced and voiceless plosives in coda position. The vowel duration from the following voiced words were analyzed:/b\ae d, bæg, k\ae b, k\textturnv b , d\textopeno g , f\ae d, fid, p\textsci g, t\ae b/. The following voiceless words were analyzed: / b\ae t, b\ae k, k\ae p, k\textturnv p, d\textscripta k, f\ae t, fit, p\textsci k, t\ae b/. Only 'post-training' productions were analyzed. For the wait-list control particpants, this was the second time that they produced these words (e.g. they did receive training in between the first and second times that they produced these words). However for trained participants, these productions represent the second time that they produced these words after they received intensive training. 

## Procedure
Participants in Terry Au's (ms) study participated in a 4 - 6 week training program compromised of comprehending and producing English phonological minmial word pairs. Not all of the words that were used in training were used in production. See _Appendix A_ for full list of words, as well as which words were used in training, and which were not. The productions were then sent to our lab for acoustic analyses.  

The software PRAAT was used to conduct acoustic analyses. Textgrids were created from the sound file in order to mark the beginning and end of the vowel boundary. Utilizing Sennheiser HD 555 headphones, the beginning of the vowel was marked with the _wav_ method and the end of the vowel was marked with the _F2_ method. All boundaries were marked at the zero-crossing line. Measurements at present, were only taken by one researcher. Thus, future cross-validation through concordance rates is required. PRAAT scripting was then used to export vowel duration measurements.  

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.
Data from the production task were analyzed using a general linear mixed-effects model using the lme4 package (1.1-10 in R 3.2.2). The criterion variable was _vowel duration_ which was convereted to milliseconds and normalized for speaker. There were two predictors which were fixed factors: (1) training _trained/untrained_ and voicing _voiced/unvoiced_. Both factors were cateogrical and were sum coded. For the training variable, _trained_ was assigned a 1, and _untrained_ was assigned a 0; while _voiced_ was assigned a 1 and _voiceless_ was assigned a 0. Two new columns in the data frame were generated to represent the sum variables of the training and the voicing variables. The variable participant was treated as a random effect as each participant had multiple productions (i.e. each participant produced each of the 36 voiced and voiceless words). 
Visual inspection of the Q-Q plots and plots of residuals against fitted values revealed that the assumptions of normality and homoscedasticity were in tact. 


# Discussion


\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
